Synthesizing Symmetric Lenses

### 2018-07-08 papers/SymmetricLenses/main.tex
add impl revisions

To obtain SREs from the plain regular expressions that
we expect users to supply, we use a heurstic that attempts to assign probability
annotations that give each string in the language equal probablity. We can also
take as input a stochastic regular expression rather than a plain one, which
allows the user to specify a data-specific probability distribution.
Alternatively, the user could provide a large suite of example data from which
we could infer the necessary annotations to model the supplied data. We found
our heuristically chosen distribution works well in practice.

@@

\begin{theorem}
  If $\Regex \SSREquiv \RegexAlt$ then $\ProbabilityOf{\Regex}{\String} =
  \ProbabilityOf{\RegexAlt}{\String}$, for all strings $\String \in \LanguageOf{\Regex}$.
\end{theorem}
\begin{proof}
  The proofs of this theorem, and subsequent theorems in this paper, appears in
  the auxiliary materials accompanying this submission.\bcp{Not in the appendix?}
\end{proof}

\subsection{Stochastic Regular Expression Entropy}
The entropy of a data source $S$ is the expected number of bits required to
describe a random element in $S$ ($-\Sigma_{s\in S}P(s)\Log_2P(s)$). The entropy
of a stochastic regular expression can be computed directly from its syntax,
when the stochastic regular expression is unambiguous and contains no nonempty\bcp{empty?}
subcomponents.
\begin{center}
  \begin{tabular}{rcl}
    $\EntropyOf{s}$
    & =
    & 1\\

@@

    $\PutL{} \App \StringAlt \App \String$
    & =
    & $\begin{cases*}
        \Lens_1.\PutLOf{\StringAlt}{\String} & if $\String\in\LanguageOf{\BRegex_1}$\\
        \Lens_2.\PutLOf{\StringAlt}{\String} & if $\String\in\LanguageOf{\BRegex_2}$
      \end{cases*}$\\
  \end{tabular}
\end{center}
The \MergeR lens is interesting because it merges data where one data can be in
two formats, and one data has only one format. In previous work\bcp{where?},
this was
combined into \OrLens{}, where \OrLens{} could have ambiguous types, but we
find it more clear to have explicit merge operators: it is easier to see what
lens the synthesis algorithm is creating.

@@

\begin{algorithm}
  \caption{\SynthSymLens}
  \label{alg:synth-sym-lens}
  \begin{algorithmic}[1]
    \Function{SynthSymLens}{$\Regex,\RegexAlt,\Examples$}
    \State $\Regex \gets \Call{\ToStochastic}{\Regex}$
    \State $\RegexAlt \gets \Call{\ToStochastic}{\RegexAlt}$
    \State $\RXSearchState \gets \Call{\RXSearch.Init}{\Regex,\RegexAlt}$
    \While{$\Call{\RXSearch.Continue}{\RXSearchState}$}
    \State $(\Regex,\RegexAlt,\RXSearchState) \gets \Call{\RXSearch.Propose}{\RXSearchState}$
    \State $\LCO \gets \Call{\GreedySynth}{\Examples,\Regex,\RegexAlt}$
    \State $\RXSearchState \gets \Call{\RXSearch.Consider}{\RXSearchState,\LCO}$
    \EndWhile
    \ReturnVal{\Call{RXSearch.Best}{\RXSearchState}}
    \EndFunction
  \end{algorithmic}
\bcp{Why does the RETURN statement not get a number?}
\end{algorithm}

@@

\RXSearch proposes rewrites incrementally, where it proposes regular expression
pairs with the smallest number of rewrites first, and when there are no more
proposed regular expressions of that size, \RXSearch proposes regular
expressions of larger size.

\RXSearch terminates when the sum of the number of regular expressions of the
smallest unproposed size with the distance of those regular expressions from the
originals exceeds the cost of the current best lens. At this point, \RXSearch
stops proposing regular expression pairs, and instead returns to the user the
best lens found thus far. Note this means that if \RXSearch finds a bijective
lens (a lens of cost 0\bcp{have we proved that the bijective lenses are
  exactly the ones with zero cost?  That would be nice to state clearly someplace...}) it will immediately return.
%
\subsection{Stochastic DNF Regular Expressions}
Stochastic DNF regular expressions are a highly normalized form of stochastic
regular expressions. Due to this highly normalized nature, where type
equivalence is often needed for simple symmetric lenses, such rules aren't
needed when typed between DNF regular expressions, as many semantically equal
but syntactically distinct regular expressions have syntactically equal DNF
forms.  Like how stochastic regular expressions are regular expressions with
probability annotations, stochastic DNF regular expressions are DNF regular
expressions with probability annotations.  Like with SREs and regular
expressions, if $\DNFRegex$ is a stochastic DNF regular expression,
$\overline{\DNFRegex}$ is the DNF regular expression generated by erasing
probability annotations.

@@

\paragraph*{\ToStochastic} With stochastic DNF regular expressions and
\ToDNFRegex defined, it is easier to explain the function that converts regular
expressions into stochastic regular expressions, \ToStochastic\bcp{from the
  description above, I thought this step was optional (because we allowed
  the user to write SREs directly)}. If $\Regex$ is a
stochastic DNF regular expression generated by \ToStochastic, then
$\ToDNFRegexOf{\Regex} = \DNFOf{(\Sequence_1,\frac{1}{n}) \DNFSep \ldots \DNFSep
  (\Sequence_n,\frac{1}{n})}$ for some sequences $\Sequence_1\ldots\Sequence_n$,
and every stochastic atom generated by \ToStochastic is
$\PRegexStar{\DNFRegex}{.8}$. In particular, $\ToDNFRegex$ generates regular
expressions whose DNF form gives equal probability to all sequences\bcp{Why?}\bcp{Also, what does this mean?  If a given RE denotes an
infinite language, then there is no way to give equal probability to every
string in that language, is there???}, and gives
iterations a .8 chance to continue iterating. In our experience generating
random strings from regular expressions\bcp{what does generating random
  strings have to do with synthesis? i.e., why is this a good thing to measure?}, these probabilities provide good
distributions of strings -- stars are iterated 4 times on average, and no
individual choice in a series disjunctions is chosen disproportionately often.

@@

\paragraph*{Syntax}
Like stochastic DNF regular expressions, symmetric DNF Lenses are hierarchically
structured. Symmetric DNF lenses consist of a list of symmetric sequence lenses,
as well as some additional information. Symmetric sequence lenses consist of a
list of symmetric atom lenses\bcp{why are they called atom lenses?}, as well
as some additional information, and
symmetric atom lenses are merely iterations of symmetric DNF lenses. In the
following syntax, the nonterminals $i$, $j$, $p$, $q$, $r$, $c$, and $d$ all
represent natural numbers. The nonterminals $\String$ and $\StringAlt$ represent
strings.
\begin{center}
  \begin{tabular}{@{}r@{\ }c@{}l@{}}
    % REGEX
    \SAtomLens{} & \GEq{} & $\StarOf{\SDNFLens}$ \\
    \SSQLens{} & \GEq{} & $(\SSQLensOf{(i_1,j_1,\SAtomLens_1)\SeqLSep
                          \ldots\SeqLSep
                          (i_p,j_p,\SAtomLens_p)}
                          ,\ListOf{(\String_1,\Atom_1);\ldots;(\String_q,\Atom_q)}
                          ,\ListOf{\StringAlt_1;\ldots;\StringAlt_r})$ \\
    \SDNFLens{} & \GEq{} & $(\SDNFLensOf{(i_1,j_1,\SSQLens_1)\DNFLSep
                           \ldots\DNFLSep
                           (i_p,j_p,\SSQLens_p)}
                           ,\ListOf{c_1;\ldots;c_q}
                           ,\ListOf{d_1;\ldots;d_r})$ \\
  \end{tabular}
\end{center}

@@

\[\begin{array}{rl}
      & \REntropyOf{\PRegexStar{\DNFRegex}{\Probability} \Given \StarOf{\DNFRegexAlt},\StarOf{\SDNFLens}}\\
      =\\
      & 4*\REntropyOf{\DNFRegex\Given\DNFRegexAlt,\SDNFLens}
  \end{array}
\]
The expected number of bits required to recover a string in $\DNFRegexAlt$ from
a synchronized string in $\DNFRegex$, $\REntropyOf{\DNFRegexAlt \Given
  \SDNFLens, \DNFRegex}$ is defined symmetrically.

\subsection{\GreedySynth}
\label{subsec:greedy-synth}
\bcp{The following is all pretty incomprehensibly dense.  A little running example
might help.}

The greedy algorithm is actually two greedy algorithms, one that greedily finds
symmetric DNF lenses (\GreedySynth) and one that greedily finds symmetric
sequence lenses (\GreedySeqSynth).

\paragraph*{Symmetric DNF Lenses} Algorithm~\ref{alg:greedysynth} presents
\GreedySynth, a means to synthesize symmetric DNF lenses. The inputs to
\GreedySynth are a pair of stochastic DNF regular expressions and a suite of
input-output examples. First, \PCF{CannotMap} checks if there could possibly be
a lens satisfying the examples. If there is not, \GreedySynth returns \None
early. After this, \GreedySynth finds the best lenses between each sequence pair
of the left and right DNF regular expressions, given the examples that match
them. An internal state for the algorithm is initialized from these. This state
consists of two things: an intermediate partial DNF Lens and a priority queue
consisting of potential sequence lenses. On initialization, \GreedySynth checks
which sequence lenses are \emph{required} to be included, from the examples.
Those are added into the partial lens, and the rest are thrown into the priority
queue. For example, an example if an input output example is provided, the
sequences that match those provided strings must be mapped to each other with a
sequence lens. \GreedySynth then loops until there are no more sequence lenses
in the priority queue. Within this loop, sequence lenses are popped. If a popped
sequence lens maps between two sequences which aren't yet included in a sequence
lens, then this sequence lens is added to the partial DNF lens. Furthermore, we
update the priorities of our sequence lenses as the algorithm proceeds: if two
sequence lenses have the same source, the second one to be popped gets a higher
cost than it originally had, to address the information that needs to be stored
for including that source of non-bijectivity. Finally, after all sequences have
been popped, the partial DNF lens is converted into a symmetric DNF lens. This
is only possible if all sequences are involved in some sequence lens: if they
are not, \PCF{GreedyState.ToDNFLens} instead returns \None.

\begin{algorithm}
  \caption{\GreedySynth}
  \label{alg:greedysynth}
  \begin{algorithmic}[1]
    \Function{GreedySynth}{$\Examples,\DNFOf{(\Sequence_1,\Probability_1)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n)},\DNFOf{(\SequenceAlt_1,\ProbabilityAlt_1)\DNFSep\ldots\DNFSep(\SequenceAlt_m,\ProbabilityAlt_m)}$}
    \If{$\Call{CannotMap}{\Examples,\DNFOf{(\Sequence_1,\Probability_1)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n)},\DNFOf{(\SequenceAlt_1,\ProbabilityAlt_1)\DNFSep\ldots\DNFSep(\SequenceAlt_m,\ProbabilityAlt_m)}}$}
    \State $\ReturnVal{\None}$
    \EndIf
    \State $\SLS \gets
    \Call{\CartesianMap}{\GreedySeqSynth(\Examples),\ListOf{\Sequence_1;\ldots;\Sequence_n},\ListOf{\SequenceAlt_1;\ldots;\SequenceAlt_m}}$
    \State $\GreedyState \gets \Call{GreedyState.Init}{\SLS}$
    \While{$\Call{GreedyState.Continue}{\GreedyState}$}
    \State $(\GreedyState,\SSQLens) \gets
    \Call{GreedyState.PopBest}{\GreedyState}$
    \If{$\Call{GreedyState.UsefulAdd}{\GreedyState,\SSQLens}$}
    \State $\GreedyState \gets \Call{GreedyState.AddSeq}{\GreedyState,\SSQLens}$
    \EndIf
    \EndWhile
    \ReturnVal{\Call{GreedyState.ToDNFLens}{\GreedyState}}
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\paragraph*{Symmetric Sequence lenses} Algorithm~\ref{alg:greedyseqsynth}
presents \GreedySeqSynth, a means to synthesizes symmetric sequence lenses. This
algorithm uses \PCF{AtomSynth}, a simple function that synthesizes atom lenses
by iterating the DNF lens between its subcomponents. The inputs to
\GreedySeqSynth are a pair of stochastic atoms and a suite of input-output
examples. First, \PCF{CannotMap} checks if there is a possibility of a lens
between the given sequences regular expressions that is consistent with the
examples. If there is not, \GreedySeqSynth returns \None early. After this,
\GreedySeqSynth finds the best lenses between each atom pair of the left and
right sequences. An internal state for the algorithm is initialized from these.
This state consists of two things: an intermediate sequence lens and a priority
queue consisting of potential sequence lenses. On initialization,
\GreedySeqSynth throws all the atom lenses into the priority queue, and the
intermediate sequence lens is one that projects all the atoms.

\GreedySeqSynth then loops until there are no more atom lenses in the priority
queue. Within this loop, atom lenses are popped. The atom lens is then added if
it is either necessary, or useful. Within a sequence lens, each atom can either
be projected or be synchronized with exactly one atom in the other sequence, and
examples can show that certain atoms \emph{need} to be mapped. In such
situations, that atom lens is necessary. A useful atom lens is where adding that
lens reduces the overall cost of the partial lens. If both atoms are not
involved in an atom lens, the lens will always be useful, but sometimes adding
that lens requires removing another lens. If the cost added by removing that
other lens is more than is removed by adding the proposed lens, then this atom
lens is not included in the sequence lens. Finally, after all atoms have been
popped, the sequence lens is validated and returned. The sequence lens is
validated if all atoms that need to be mapped, are: if they are not,
\PCF{GreedySeqState.ToDNFLens} instead returns \None.

@@

    \Function{GreedySeqSynth}{$\Examples,\SequenceOf{\String_0\SeqSep\Atom_1\SeqSep\ldots\SeqSep\Atom_n\SeqSep\String_n},\SequenceOf{\StringAlt_0\SeqSep\AtomAlt_1\SeqSep\ldots\SeqSep\AtomAlt_m\SeqSep\String_m}$}
    \If{$\Call{CannotMap}{\Examples,\SequenceOf{\String_0\SeqSep\Atom_1\SeqSep\ldots\SeqSep\Atom_n\SeqSep\String_n},\SequenceOf{\StringAlt_0\SeqSep\AtomAlt_1\SeqSep\ldots\SeqSep\AtomAlt_m\SeqSep\String_m}}$}
    \State $\ReturnVal{\None}$
    \EndIf
    \State $\ALS \gets
    \Call{\CartesianMap}{\AtomSynth(\Examples),\ListOf{\Atom_1;\ldots;\Atom_n},\ListOf{\AtomAlt_1;\ldots;\AtomAlt_m}}$
    \State $\GreedyState \gets \Call{GreedySeqState.Init}{\ALS}$
    \While{$\Call{GreedySeqState.Continue}{\GreedyState}$}
    \State $(\GreedyState,\SAtomLens) \gets
    \Call{GreedySeqState.PopBest}{\GreedyState}$
    \If{$\Call{GreedySeqState.UsefulAdd}{\GreedyState,\SAtomLens}
      \BooleanOr \Call{NecessaryAdd}{\Examples,\SAtomLens}$}
    \State $\GreedyState \gets \Call{GreedySeqState.AddAtom}{\GreedyState,\SAtomLens}$
    \EndIf
    \EndWhile
    \ReturnVal{\Call{GreedySeqState.ToDNFLens}{\GreedyState}}
    \EndFunction
  \end{algorithmic}
\end{algorithm}

@@

\paragraph*{\RXSearch Permissiveness Customization}\bcp{I don't follow this:}
While the termination heuristic works for many situations, there are situations
where a large number of expansions must be performed to find the correct lens.
While users could prevent incorrect lenses with a large number of examples, we
find it simpler to let the users specify that \RXSearch should be more
restrictive in which lenses it accepts. If the user's desired transformation has
a particularly high cost, the algorithm may spend a long time searching in vain
for lenses with lower cost.  By tuning \RXSearch to me more permissive in the
lenses it accepts, this search can be short-circuited.
% end synthesis

>>>>

To obtain SREs from the plain regular expressions that we expect users to
supply, we use a heurstic that attempts to assign probability annotations that
give each string in the language equal probablity. Our algorithm could also take
as input a stochastic regular expression rather than a plain one, which allows
the user to specify a data-specific probability distribution, either manually
written or inferred from a large dataset. We found our heuristically chosen
distribution works well in practice.

@@

\begin{theorem}
  If $\Regex \SSREquiv \RegexAlt$ then $\ProbabilityOf{\Regex}{\String} =
  \ProbabilityOf{\RegexAlt}{\String}$, for all strings $\String \in \LanguageOf{\Regex}$.
\end{theorem}
\begin{proof}
  The proofs of this theorem, and subsequent theorems in this paper, appears in
  the appendix present in the full version of the paper accompanying this
  submission.
\end{proof}

\subsection{Stochastic Regular Expression Entropy}
The entropy of a data source $S$ is the expected number of bits required to
describe a random element in $S$ ($-\Sigma_{s\in S}P(s)\Log_2P(s)$). The entropy
of a stochastic regular expression can be computed directly from its syntax,
when the stochastic regular expression is unambiguous and contains no empty
subcomponents.
\begin{center}
  \begin{tabular}{rcl}
    $\EntropyOf{s}$
    & =
    & 1\\

@@

    $\PutL{} \App \StringAlt \App \String$
    & =
    & $\begin{cases*}
        \Lens_1.\PutLOf{\StringAlt}{\String} & if $\String\in\LanguageOf{\BRegex_1}$\\
        \Lens_2.\PutLOf{\StringAlt}{\String} & if $\String\in\LanguageOf{\BRegex_2}$
      \end{cases*}$\\
  \end{tabular}
\end{center}
The \MergeR lens is interesting because it merges data where one data can be in
two formats, and one data has only one format. In previous work~\cite{?}, this
was combined into \OrLens{}, where \OrLens{} could have ambiguous types, but we
find it more clear to have explicit merge operators: it is easier to see what
lens the synthesis algorithm is creating.

@@

\begin{algorithm}
  \caption{\SynthSymLens}
  \label{alg:synth-sym-lens}
  \begin{algorithmic}[1]
    \Function{SynthSymLens}{$\Regex,\RegexAlt,\Examples$}
    \State $\Regex \gets \Call{\ToStochastic}{\Regex}$
    \State $\RegexAlt \gets \Call{\ToStochastic}{\RegexAlt}$
    \State $\RXSearchState \gets \Call{\RXSearch.Init}{\Regex,\RegexAlt}$
    \While{$\Call{\RXSearch.Continue}{\RXSearchState}$}
    \State $(\Regex,\RegexAlt,\RXSearchState) \gets \Call{\RXSearch.Propose}{\RXSearchState}$
    \State $\LCO \gets \Call{\GreedySynth}{\Examples,\Regex,\RegexAlt}$
    \State $\RXSearchState \gets \Call{\RXSearch.Consider}{\RXSearchState,\LCO}$
    \EndWhile
    \State \ReturnVal{\Call{RXSearch.Best}{\RXSearchState}}
    \EndFunction
  \end{algorithmic}
\end{algorithm}

@@

\RXSearch proposes rewrites incrementally, where it proposes regular expression
pairs with the smallest number of rewrites first. When there are no more
proposed regular expressions derived from $n$ rewrites, \RXSearch proposes
regular expressions derived from $n+1$ rewrites.

\RXSearch terminates when the sum of the number of regular expressions of the
smallest unproposed size with the distance of those regular expressions from the
originals exceeds the cost of the current best lens. At this point, \RXSearch
stops proposing regular expression pairs, and instead returns to the user the
best lens found thus far. Note this means that if \RXSearch finds a bijective
lens, which has zero cost it will immediately return.
%
\subsection{Stochastic DNF Regular Expressions}
Stochastic DNF regular expressions are a highly normalized form of stochastic
regular expressions. Due to this highly normalized nature, where type
equivalence is often needed for simple symmetric lenses, such rules aren't
needed when typed between DNF regular expressions, as many semantically equal
but syntactically distinct regular expressions have syntactically equal DNF
forms.  Like how stochastic regular expressions are regular expressions with
probability annotations, stochastic DNF regular expressions are DNF regular
expressions with probability annotations.  Like with SREs and regular
expressions, if $\DNFRegex$ is a stochastic DNF regular expression,
$\overline{\DNFRegex}$ is the DNF regular expression generated by erasing
probability annotations.

@@

\paragraph*{\ToStochastic} With stochastic DNF regular expressions and
\ToDNFRegex defined, it is easier to explain the function that converts regular
expressions into stochastic regular expressions, \ToStochastic. If users
specified stochastic regular expressions themselves, this step could be skipped,
though our integration into Boomerang does not permit users to provide
stochastic regular expressions. If $\Regex$ is a stochastic DNF regular
expression generated by \ToStochastic, then $\ToDNFRegexOf{\Regex} =
\DNFOf{(\Sequence_1,\frac{1}{n}) \DNFSep \ldots \DNFSep
  (\Sequence_n,\frac{1}{n})}$ for some sequences $\Sequence_1\ldots\Sequence_n$,
and every stochastic atom generated by \ToStochastic is
$\PRegexStar{\DNFRegex}{.8}$. In particular, $\ToDNFRegex$ generates regular
expressions whose DNF form gives equal probability to all sequence subcomponents
of the DNF regular expression., and gives iterations a .8 chance to continue
iterating. In our experience generating random strings from regular expressions,
these probabilities provide good distributions of strings -- stars are iterated
4 times on average, and no individual choice in a series disjunctions is chosen
disproportionately often. As we have found these distributions generate
reasonable strings, we believe they describe the shape of data well.

@@

\paragraph*{Syntax}
Like stochastic DNF regular expressions, symmetric DNF Lenses are hierarchically
structured. Symmetric DNF lenses, which provide lens functions between DNF
regular expressions, consist of a list of symmetric sequence lenses, as well as
some additional information. Symmetric sequence lenses, which provide lens
functions betwen sequences, consist of a list of symmetric atom lenses, as well
as some additional information. Symmetric atom lenses, which provide lens
functions between atoms, are merely iterations of symmetric DNF lenses. In the
following syntax, the nonterminals $i$, $j$, $p$, $q$, $r$, $c$, and $d$ all
represent natural numbers. The nonterminals $\String$ and $\StringAlt$ represent
strings.
\begin{center}
  \begin{tabular}{@{}r@{\ }c@{}l@{}}
    % REGEX
    \SAtomLens{} & \GEq{} & $\StarOf{\SDNFLens}$ \\
    \SSQLens{} & \GEq{} & $(\SSQLensOf{(i_1,j_1,\SAtomLens_1)\SeqLSep
                          \ldots\SeqLSep
                          (i_p,j_p,\SAtomLens_p)}
                          ,\ListOf{(\String_1,\Atom_1);\ldots;(\String_q,\Atom_q)}
                          ,\ListOf{\StringAlt_1;\ldots;\StringAlt_r})$ \\
    \SDNFLens{} & \GEq{} & $(\SDNFLensOf{(i_1,j_1,\SSQLens_1)\DNFLSep
                           \ldots\DNFLSep
                           (i_p,j_p,\SSQLens_p)}
                           ,\ListOf{c_1;\ldots;c_q}
                           ,\ListOf{d_1;\ldots;d_r})$ \\
  \end{tabular}
\end{center}

@@

\[\begin{array}{rl}
      & \REntropyOf{\PRegexStar{\DNFRegexAlt}{\ProbabilityAlt} \Given \PRegexStar{\DNFRegex}{\Probability},\StarOf{\SDNFLens}}\\
      =
      & \frac{\Probability}{1-\Probability}\REntropyOf{\DNFRegexAlt\Given\DNFRegex,\SDNFLens}
  \end{array}
\]
The expected number of bits required to recover a string in $\DNFRegexAlt$ from
a synchronized string in $\DNFRegex$, $\REntropyOf{\DNFRegexAlt \Given
  \SDNFLens, \DNFRegex}$ is defined symmetrically.

\subsection{\GreedySynth}
\label{subsec:greedy-synth}
The greedy algorithm is actually two greedy algorithms, one that greedily finds
symmetric DNF lenses (\GreedySynth) and one that greedily finds symmetric
sequence lenses (\GreedySeqSynth).
We describe this synthesis algorithm in the context of trying to find a lens
between \lstinline{"" | name.name*} and \lstinline{"" | name}, with the example
\lstinline{creater "John Doe" = "John"}.

\paragraph*{Symmetric DNF Lenses} Algorithm~\ref{alg:greedysynth} presents
\GreedySynth, a means to synthesize symmetric DNF lenses. The inputs to
\GreedySynth are a pair of stochastic DNF regular expressions and a suite of
input-output examples.

First, \PCF{CannotMap} checks if there could possibly be a lens satisfying the
examples. If there is not, \GreedySynth returns \None early. After this,
\GreedySynth finds the best lenses between each sequence pair of the left and
right DNF regular expressions, given the examples that match them. An internal
state for the algorithm is initialized from these. This state consists of two
things: an intermediate partial DNF Lens and a priority queue consisting of
potential sequence lenses.

\GreedySynth then loops until there are no more sequence lenses in the priority
queue. Within this loop, sequence lenses are popped. If a popped sequence lens
maps between two sequences which aren't yet included in a sequence lens, then
this sequence lens is added to the partial DNF lens. Furthermore, we update the
priorities of our sequence lenses as the algorithm proceeds: if two sequence
lenses have the same source, the second one to be popped gets a higher cost than
it originally had, to address the information that needs to be stored for
including that source of non-bijectivity.

In our example, \GreedySynth may first pop the sequence pair \lstinline{""} and
\lstinline{""}, as there is a bijective sequence lens between them. After this,
the sequence lens between \lstinline{name.name*} and \lstinline{name} would be
popped. While examples prevent lenses between the other sequence pairs, if there
were other lenses, they would be popped, but not added to the ongoing state, as
they would not be a useful addition to the state.

Finally, after all sequences have been
popped, the partial DNF lens is converted into a symmetric DNF lens. This is
only possible if all sequences are involved in some sequence lens: if they are
not, \PCF{GreedyState.ToDNFLens} instead returns \None.

\begin{algorithm}
  \caption{\GreedySynth}
  \label{alg:greedysynth}
  \begin{algorithmic}[1]
    \Function{GreedySynth}{$\Examples,\DNFOf{(\Sequence_1,\Probability_1)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n)},\DNFOf{(\SequenceAlt_1,\ProbabilityAlt_1)\DNFSep\ldots\DNFSep(\SequenceAlt_m,\ProbabilityAlt_m)}$}
    \If{$\Call{CannotMap}{\Examples,\DNFOf{(\Sequence_1,\Probability_1)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n)},\DNFOf{(\SequenceAlt_1,\ProbabilityAlt_1)\DNFSep\ldots\DNFSep(\SequenceAlt_m,\ProbabilityAlt_m)}}$}
    \State $\ReturnVal{\None}$
    \EndIf
    \State $\SLS \gets
    \Call{\CartesianMap}{\GreedySeqSynth(\Examples),\ListOf{\Sequence_1;\ldots;\Sequence_n},\ListOf{\SequenceAlt_1;\ldots;\SequenceAlt_m}}$
    \State $\GreedyState \gets \Call{GreedyState.Init}{\SLS}$
    \While{$\Call{GreedyState.Continue}{\GreedyState}$}
    \State $(\GreedyState,\SSQLens) \gets
    \Call{GreedyState.PopBest}{\GreedyState}$
    \If{$\Call{GreedyState.UsefulAdd}{\GreedyState,\SSQLens}$}
    \State $\GreedyState \gets \Call{GreedyState.AddSeq}{\GreedyState,\SSQLens}$
    \EndIf
    \EndWhile
    \State \ReturnVal{\Call{GreedyState.ToDNFLens}{\GreedyState}}
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\paragraph*{Symmetric Sequence lenses} Algorithm~\ref{alg:greedyseqsynth}
presents \GreedySeqSynth, a means to synthesizes symmetric sequence lenses.

This algorithm uses \PCF{AtomSynth}, a simple function that synthesizes atom
lenses by iterating the DNF lens between its subcomponents. The inputs to
\GreedySeqSynth are a pair of stochastic atoms and a suite of input-output
examples. First, \PCF{CannotMap} checks if there is a possibility of a lens
between the given sequences regular expressions that is consistent with the
examples. If there is not, \GreedySeqSynth returns \None early. After this,
\GreedySeqSynth finds the best lenses between each atom pair of the left and
right sequences. An internal state for the algorithm is initialized from these.
This state consists of two things: an intermediate sequence lens and a priority
queue consisting of potential sequence lenses. On initialization,
\GreedySeqSynth throws all the atom lenses into the priority queue, and the
intermediate sequence lens is one that projects all the atoms.

For example, if trying to find a sequence lens between \lstinline{name.name*}
and \lstinline{name}, lenses are found between \lstinline{name} and
\lstinline{name}, and \lstinline{name} and \lstinline{name*}, and added to the
queue.

\GreedySeqSynth then loops until there are no more atom lenses in the priority
queue. Within this loop, atom lenses are popped. The atom lens is then added if
it is either necessary, or useful. Within a sequence lens, each atom can either
be projected or be synchronized with exactly one atom in the other sequence, and
examples can show that certain atoms \emph{need} to be mapped. In such
situations, that atom lens is necessary. A useful atom lens is where adding that
lens reduces the overall cost of the partial lens. If both atoms are not
involved in an atom lens, the lens will always be useful, but sometimes adding
that lens requires removing another lens. If the cost added by removing that
other lens is more than is removed by adding the proposed lens, then this atom
lens is not included in the sequence lens. For example, the bijective lens
between \lstinline{name} and \lstinline{name} would first be found, and added to
the state.  Next, the disconnect lens between \lstinline{name} and
\lstinline{name*} would be popped.  Because this lens is not useful, it is not
added to the state.

Finally, after all atoms have been
popped, the sequence lens is validated and returned. The sequence lens is
validated if all atoms that need to be mapped, are: if they are not,
\PCF{GreedySeqState.ToDNFLens} instead returns \None.

@@

    \Function{GreedySeqSynth}{$\Examples,\SequenceOf{\String_0\SeqSep\Atom_1\SeqSep\ldots\SeqSep\Atom_n\SeqSep\String_n},\SequenceOf{\StringAlt_0\SeqSep\AtomAlt_1\SeqSep\ldots\SeqSep\AtomAlt_m\SeqSep\String_m}$}
    \If{$\Call{CannotMap}{\Examples,\SequenceOf{\String_0\SeqSep\Atom_1\SeqSep\ldots\SeqSep\Atom_n\SeqSep\String_n},\SequenceOf{\StringAlt_0\SeqSep\AtomAlt_1\SeqSep\ldots\SeqSep\AtomAlt_m\SeqSep\String_m}}$}
    \State $\ReturnVal{\None}$
    \EndIf
    \State $\ALS \gets
    \Call{\CartesianMap}{\AtomSynth(\Examples),\ListOf{\Atom_1;\ldots;\Atom_n},\ListOf{\AtomAlt_1;\ldots;\AtomAlt_m}}$
    \State $\GreedyState \gets \Call{GreedySeqState.Init}{\ALS}$
    \While{$\Call{GreedySeqState.Continue}{\GreedyState}$}
    \State $(\GreedyState,\SAtomLens) \gets
    \Call{GreedySeqState.PopBest}{\GreedyState}$
    \If{$\Call{GreedySeqState.UsefulAdd}{\GreedyState,\SAtomLens}
      \BooleanOr \Call{NecessaryAdd}{\Examples,\SAtomLens}$}
    \State $\GreedyState \gets \Call{GreedySeqState.AddAtom}{\GreedyState,\SAtomLens}$
    \EndIf
    \EndWhile
    \State \ReturnVal{\Call{GreedySeqState.ToDNFLens}{\GreedyState}}
    \EndFunction
  \end{algorithmic}
\end{algorithm}

@@

\paragraph*{\RXSearch Permissiveness Customization}
In this work, we are constrained by three competing constraints. (1) We want to
return lenses quickly to users, (2) we want to return good lenses, and (3) we
don't want to require users to write a large number of examples in the
specifications. Our termination heuristic works well for most situations, but
there are situations where a large number of expansions must be performed to
find the correct lens.

Users could prevent incorrect lenses with a large number of examples, but it can
be quite difficult to write a large number of examples for complex formats. We
find it simpler to let the users specify that \RXSearch should be more
restrictive in which lenses it accepts, so users do not have to write a lot of
examples.

This constraint goes the other way too. If the user's desired transformation has
a particularly high cost, the algorithm may spend a long time searching in vain
for lenses with lower cost. By tuning \RXSearch to me more permissive in the
lenses it accepts, this search can be short-circuited.

While our defaults termination condition quickly finds the correct lens for
\afm{35} of our \afm{45} benchmarks, our termination condition is not
one-size-fits all. Particularly complex lenses, or lenses that project a
particularly large amount of data, can be synthesized more easily and more
quickly by customizing how permissive \RXSearch is in returning lenses.
% end synthesis
